### 2017-03-01 paper/chapters/vae.tex
fixed weird formulation

% more math:
% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
To be able to apply the backpropagation\cite{backprop:1988} algorithm to the loss function of the VAE without any injected noise it has to be both differentiable and deterministic.
However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
In order to circumvent this restriction, the so-called ``reparameterization trick'' (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) is applied.
Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$. Then we apply Equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
\begin{equation}
  \label{eq:rep_trick}
  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
\end{equation}
Note that this transformation is entirely deterministic and as such allows for backpropagation through the transformation into the variational parameters $\phi$.

>>>>

% more math:
% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
To be able to apply the backpropagation\cite{backprop:1988} algorithm to the loss function of the VAE without any injected noise it has to be both differentiable and deterministic.
However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
In order to circumvent this restriction, the reparameterization trick (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) can be applied.
Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$. Then we apply Equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
\begin{equation}
  \label{eq:rep_trick}
  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
\end{equation}
Note that this transformation is entirely deterministic and as such allows for backpropagation through the transformation into the variational parameters $\phi$.
